\section{Introduction}

\newcommand{\mandragore}{\textcolor{blue}{mandragore}}
\newcommand{\vulpecula}{\textcolor{green}{vulpecula}}
\newcommand{\noscope}{\textcolor{red}{noscope}}

\begin{frame}
  \frametitle{What's the buzz with multi-party Instant Messaging?}
  \centering
  Instant messaging with two participants is solved!
  Why are group chats not encrypted already?!\\[0.3cm]
\end{frame}

\begin{frame}
  \centering
  Answer:

  Because we still can't solve every problem at once! So we have to choose.
\end{frame}

\begin{frame}

  Here are some problems to choose from:

  \begin{itemize}
    \item Chatroom Message Consistency
    \item Room Moderation
    \item Offline messages but with a group secret
  \end{itemize}

  \vfill

  We can't solve all these problems together. But we can solve some of them.

\end{frame}

\begin{frame}
  Our implementation is based on the original mpOTR pape and this is what it provides:

  \begin{itemize}
    \item Confidentiality
    \item Authentication
    \item Forward Secrecy
    \item Deniable participation in chatroom
    \item Deniable authoring of messages
    \item Partial Transcript Consistency
  \end{itemize}

\end{frame}

\begin{frame}
These properties are well known in the crypto community.
\vfill
Except for Transcript Consistency!
\vfill
So let's have a word about it.
\end{frame}

\begin{frame}
  \frametitle{Who says no to icecream?}
    \begin{itemize}
      \item[] \mandragore: Who wants icecream?
      \item[] \noscope: I do!
      \item[] \mandragore: Who wants to do something totally illegal?
      \item[] \vulpecula: what the hell?
    \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Who says no to icecream?}
    \begin{itemize}
      \item[] \mandragore: Who wants to do something totally illegal?
      \item[] \noscope: I do!
      \item[] \vulpecula: what the hell?
    \end{itemize}
\end{frame}

\begin{frame}
  And now let's see what we did.
\end{frame}
